---
title: "Analysis of Response Data"
author: "Semih Can Aktepe"
date: "2023-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### First set the working directory.
All the data files are in this folder, so we extract the data from here.
```{r}
setwd("/Users/semih/Desktop/cf_responses")
```

### Load the necessary packages
```{r}
library(dplyr)
library(ggplot2)
library(brms)
library(posterior)
library(BayesFactor)
library(bayestestR)
```

### Load the data set
```{r}
response <- read.csv("~/Desktop/response.csv")
```

## Preparation

### Convert the data types into appropriate formats
```{r}
response$id <- as.factor(response$id)
response$subject <- as.factor(response$subject)
response$group <- as.factor(response$group)
response$item <- as.factor(response$item)
response$condition <- as.factor(response$condition)
response$outcome <- as.factor(response$outcome)
```

### Demographic information
```{r}
# Age
age <- aggregate(age ~ group + subject, data = response, FUN = max) %>% arrange(group)

# Adults age mean, standard deviation, minimum and maximum
n_ad <- 17
n_ad_fem <- 10
age_ad <- mean(age[age$group == "Adult",]$age, na.rm = TRUE)
sd_ad <- sd(age[age$group == "Adult",]$age, na.rm = TRUE)
min_ad <- min(age[age$group == "Adult",]$age, na.rm = TRUE)
max_ad <- max(age[age$group == "Adult",]$age, na.rm = TRUE)
adult_demo <- c(n_ad, age_ad, sd_ad, min_ad, max_ad)

# Children age mean, standard deviation, minimum and maximum
n_ch <- 39-n_ad
n_ch_fem <- 12
age_ch <- mean(age[age$group == "Child",]$age, na.rm = TRUE)/12
sd_ch <- sd(age[age$group == "Child",]$age, na.rm = TRUE)/12
min_ch <- min(age[age$group == "Child",]$age, na.rm = TRUE)/12
max_ch <- max(age[age$group == "Child",]$age, na.rm = TRUE)/12
child_demo <- c(n_ch, age_ch, sd_ch, min_ch, max_ch)

# Print the results
adult_demo %>% round(2)
child_demo %>% round(2)
```

### Visualization

### Data preparation
```{r}
response$ss <- 1 # Dummy variable for standard error and confidence interval calculations.
d1 <- aggregate(accuracy ~ group + condition + outcome, data = response, FUN = mean)
d2 <- aggregate(accuracy ~ group + condition + outcome, data = response, FUN = sd)
d3 <- aggregate(ss ~ group + condition + outcome, data = response, FUN = sum)

d <- cbind(d1, sd = d2$accuracy, ss = d3$ss) %>% arrange(condition, outcome)
d$se <- d$sd/sqrt(d$ss)
alpha = 0.05
d$ci <- qt(p=alpha/2, df=d$ss,lower.tail=F)*(d$sd/sqrt(d$ss))
d$Low <- d$accuracy - d$ci
d$High <- d$accuracy + d$ci
rm(d1,d2,d3)
```

### Define a new variable for ease
```{r}
d$itemtype <- "Counterfactual Positive"
d$itemtype[d$condition == "Counterfactual" & d$outcome == "Negative"] <- "Counterfactual Negative"
d$itemtype[d$condition == "Indicative" & d$outcome == "Positive"] <- "Indicative Positive"
d$itemtype[d$condition == "Indicative" & d$outcome == "Negative"] <- "Indicative Negative"
```

### Define custom colors
```{r}
col1 <- rgb(red = 1, green = 0.647, blue = 0, alpha = 1)
col1l <- rgb(red = 1, green = 0.647, blue = 0, alpha = 0.025)
col2 <- rgb(red = 0.12, green = 0.56, blue = 1, alpha = 1)
col2l <- rgb(red = 0.12, green = 0.56, blue = 1, alpha = 0.025)
col3 <- rgb(red = 0.196, green = 0.804, blue = 0.196, alpha = 1)
col3l <- rgb(red = 0.196, green = 0.804, blue = 0.196, alpha = 0.025)
col4 <- rgb(red = 0.969, green = 0.325, blue = 0.58, alpha = 1)
col4l <- rgb(red = 0.969, green = 0.325, blue = 0.58, alpha = 0.025)
```

### Barplots with 95% Confidence Intervals
```{r}
ggplot(data = d, aes(x = itemtype, y = accuracy*100, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = (accuracy - se)*100, ymax = (accuracy + se)*100),
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(title = "Percentage of Correct Responses (95% CI)",
       x = "Type of Conditional Sentence", y = "Accuracy (%)",
       fill = "Age Group") +
  scale_fill_manual(labels=c("Adults", "Children"), values=c(col1, col2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,105)) +
  theme_classic()
```

## Statistical Analysis

### Sum Coding
```{r}
response$Group <- ifelse(response$group == "Adult", 1, -1)
response$Condition <- ifelse(response$condition == "Counterfactual", 1, -1)
response$Outcome <- ifelse(response$outcome == "Negative", 1, -1)
```

### Fit the model
```{r}
m1 <- brm(accuracy ~ 1 + Group*Condition*Outcome # Fixed Effects
          + (1 + Group*Condition*Outcome | subject) # Varying intercept and slope of subject
          + (1 + Group*Condition*Outcome | item), # Varying intercept and slope of item
          prior = c(prior(normal(0, 1), class = Intercept),
                    prior(normal(0, 1), class = b),
                    prior(exponential(1), class = sd),
                    prior(lkj(4), class = cor)),
          data = response, family = bernoulli, sample_prior = TRUE,
          iter = 4000, warmup = 2000, chains = 4, cores = 8,
          control = list(adapt_delta = 0.99, max_treedepth = 12))
```

### Check the model fit
```{r}
plot(m1, ask = FALSE)
```

### Posterior Predictive Check
```{r}
pp_check(m1)
```

### Model Summary
```{r}
summary(m1, prob = .95)
```

### Hypothesis testing of whether the fixed effects deviate from 0
```{r}
hypotheses <- c("Group = 0", # There is no effect of group.
                "Condition = 0", # There is no effect of condition.
                "Outcome = 0") # There is no effect of outcome
hyp <- hypothesis(m1, hypothesis = hypotheses, class = "b")
plot(hyp)
```

```{r}
hyp$hypothesis$Evid.Ratio %>% round(digits = 2)

# H0: There is no effect of group, condition, and outcome, respectively.
# H1: There is an effect of group, condition, and outcome, respectively.

# Group: Extreme evidence for H1
# Condition: Anecdotal evidence for H0
# Outcome: Anecdotal evidence for H1
```

### Conditional effects
```{r}
plot(conditional_effects(m1), ask = FALSE)
```

### Leave-One-Out (LOO) Cross Validation
```{r}
loo(m1)
```

### Print Bayesian R^2
```{r}
bayes_R2(m1)
```

### Extract the posterior samples of the parameters.
```{r}
post <- as_draws_df(m1)
```

### If an effect is negative, check the probability that it is positive; if positive, vice versa.
```{r}
mean(post$b_Group < 0)
mean(post$b_Condition > 0)
mean(post$b_Outcome > 0)
mean(post$`b_Group:Condition` < 0)
mean(post$`b_Group:Outcome` < 0)
mean(post$`b_Condition:Outcome` > 0)
mean(post$`b_Group:Condition:Outcome` > 0)
```

### Posterior Predictive Simulation for the Fixed Effects

### Prepare the simulation data
```{r}
group <- rep(c("Adult", "Child"), times = 4)
condition <- rep(c("Counterfactual", "Indicative"), each = 4)
outcome <- rep(rep(c("Negative", "Positive"), each = 2), times = 2)
newdata <- data.frame(group = group, condition = condition, outcome = outcome)
newdata$Group <- ifelse(newdata$group == "Adult", 1, -1)
newdata$Condition <- ifelse(newdata$condition == "Counterfactual", 1, -1)
newdata$Outcome <- ifelse(newdata$outcome == "Negative", 1, -1)
```

### Posterior Predictives excluding random effects
```{r}
pp <- as.data.frame(posterior_epred(m1, newdata = newdata, re_formula = NA))
pp2 <- ci(pp, ci=0.95, method = "HDI")
```

### Visualization of Posterior Predictives
```{r}
# Define custom points for where to plot the data
cp <- c(0.1, 0.15, 0.25, 0.3, 0.4, 0.45, 0.55, 0.6)
# Define the labels to replace numeric points
labels <- c("Counterfactual\nNegative", "Counterfactual\nPositive", 
            "Indicative\nNegative", "Indicative\nPositive")
# Plot the data
plot(NULL, xlim = c(0.05,0.65), ylim = c(0,100), 
     main = "Posterior Predictive Distribution of Accuracy (95% CrI)",
     xlab = "Type of Conditional Sentence", ylab = "Accuracy (%)", xaxt = "n", bty = "l")
for (i in 1:8){
  for (j in 1:8000){
    points(cp[i], pp[j,i]*100, pch = 19, cex = 1, col = ifelse(i%%2==1, col1l, col2l))
    arrows(cp[i], pp2$CI_low[i]*100, cp[i], pp2$CI_high[i]*100, length = 0, code=3, col = "black", lwd = 1)
    points(cp[i], mean(pp[,i])*100, pch = 21, cex = 0.75, col = "black", 
           bg = ifelse(i%%2==1, col1, col2))
  }
}
axis(1, at = c(0.125, 0.275, 0.425, 0.575), labels = labels, cex.axis = 0.75)
legend("bottomright", inset = 0.03, title = "Age Group",  c("Adults", "Children"),
       col = c(col1, col2), pch = 19, cex = 0.75)
```

### Posterior Predictive Simulation for Varying Effect of Subject

### Prepare the simulation data
```{r}
subject <- rep(1:39, each = 4)
condition <- rep(rep(c("Counterfactual", "Indicative"), each = 2), times = 39)
outcome <- rep(c("Negative", "Positive"), times = 39*2)
group <- c(rep("Adult", times = 16), rep("Child", times = 39-16))
newdataS <- data.frame(subject = subject, group = group, condition = condition, outcome = outcome)
newdataS$Group <- ifelse(newdataS$group == "Adult", 1, -1)
newdataS$Condition <- ifelse(newdataS$condition == "Counterfactual", 1, -1)
newdataS$Outcome <- ifelse(newdataS$outcome == "Negative", 1, -1)
```

### Map the simulations to the new data.
```{r}
pps <- as.data.frame(posterior_epred(m1, newdata = newdataS, 
                                     re_formula = ~ (1 + Group*Condition*Outcome | subject)))
pps2 <- ci(pps, method = "HDI")
```

### Visualization
```{r}
cps <- 1:156
random_sample <- sample(1:8000, 1000, replace = TRUE)
# Plot the data
plot(NULL, xlim = c(1,156), ylim = c(0,100), 
     main = "Posterior Predictive Distribution of Accuracy (95% CrI)",
     xlab = "Subject", ylab = "Accuracy (%)", xaxt = "n", bty = "l")
for (i in 1:156){
  for (j in random_sample){
    points(cps[i], pps[j,i]*100, pch = 16, cex = 0.5,
           col = ifelse(i%%4==1, col1l,
                        ifelse(i%%4==2, col2l,
                               ifelse(i%%4==3, col3l, col4l))))
    #arrows(cps[i], pps2$CI_low[i]*100, cps[i], pps2$CI_high[i]*100, length = 0, code = 3, col = "black", lwd = 0.33)
    points(cps[i], mean(pps[,i])*100, pch = 21, cex = 0.5, col = "black", 
           bg = ifelse(i%%4==1, col1,
                       ifelse(i%%4==2, col2,
                              ifelse(i%%4==3, col3, col4))))
  }
}
abline(v = 68.5, lty = 3, lwd = 0.75)
axis(1, at = seq(from=2.5, to = 153.5, length.out = 39), labels = 1:39, cex.axis = 0.75)
legend("bottomleft", inset = 0.01, title = "Type", c("CN", "CP", "IN", "IP"),
       col = c(col1, col2, col3, col4), pch = 16, cex = 0.75)
```

## Effect of Executive Functions on Accuracy

### Z-transformation of the variables
```{r}
response$zI <- scale(response$inhibition)[,1]
response$zSTM <- scale(response$stm)[,1]
response$zWM <- scale(response$wm)[,1]
```

### Aggregation for visualization
```{r}
ef0 <- aggregate(accuracy ~ subject, data = response, FUN = sum)
ef0$zA <- scale(ef0$accuracy)[,1]
ef0$subject <- as.integer(ef0$subject)
ef1 <- aggregate(zI ~ subject, data = response, FUN = max)
ef1$subject <- as.integer(ef1$subject)
ef2 <- aggregate(zSTM ~ subject, data = response, FUN = max)
ef2$subject <- as.integer(ef2$subject)
ef3 <- aggregate(zWM ~ subject, data = response, FUN = max)
ef3$subject <- as.integer(ef3$subject)
```

### Visualization
```{r}
plot(NULL, xlim = c(1,39), ylim = c(-3,3), xlab = "Subject", ylab = "Score (z-transformed)",
     main = "Distribution of Cognitive Task Scores", bty = "l", xaxt = "n")
points(ef0$subject, ef0$zA, col = "black", pch = 8, cex = 0.75)
points(ef1$subject, ef1$zI, col = col1, pch = 19)
points(ef2$subject, ef2$zSTM, col = col2, pch = 15)
points(ef3$subject, ef3$zWM, col = col3, pch = 17)
abline(v=16.5, lty = 3, lwd = 1)
axis(1, at = 1:39)
text(c(8, 25), c(-2,2), c("Adults", "Children"), font = 2)
legend("topright", inset = 0.03, title = "Task", c("Experiment", "Inhibition", "STM", "WM"),
       col = c("black", col1, col2, col3), pch = c(8,19,15,17), cex = 0.7)
```

## Analysis of Children Only

### Subset the data to analyze only the children
```{r}
children <- subset(response, group == "Child")
children$zA <- scale(children$age)[,1]
```

### Bayesian Correlations for Children Only
```{r}
# Extract the necessary columns
corDat <- data.frame(age=children$zA,
                     acc=children$accuracy,
                     ihb=children$zI,
                     stm=children$zSTM,
                     wm=children$zWM)

# Create a correlation table
corrTable <- as.data.frame(matrix(nrow = 5, ncol = 5))
rownames(corrTable) <- c("Age", "Accuracy", "Inhibition", "Short-Term", "Working")
colnames(corrTable) <- c("Age", "Accuracy", "Inhibition", "Short-Term", "Working")

# Compute posterior rho's
for (i in 1:5){
  for (j in 1:5){
    sample <- correlationBF(y = corDat[,i], x = corDat[,j],
                            posterior = TRUE, iterations = 1e4,
                            rscale = 1)
    corrTable[i,j] <- paste(round(mean(sample[,"rho"]), 2),
                            " CrI: [",round(ci(as.vector(sample[,"rho"]), method = "HDI")$CI_low, 2),
                            ", ", round(ci(as.vector(sample[,"rho"]), method = "HDI")$CI_high, 2), "]",
                            sep = "")
  }
}

# Print the correlation table
corrTable
```

#### Fit the model
```{r}
m2 <- brm(accuracy ~ 1 + zA + zI + zSTM + zWM # Fixed Effects
          + (1 + zA + zI + zSTM + zWM | subject) # Varying intercept and slope of subject
          + (1 + zA + zI + zSTM + zWM | item), # Varying intercept and slope of item
          prior = c(prior(normal(0, 1), class = Intercept),
                    prior(normal(0, 1), class = b),
                    prior(exponential(1), class = sd),
                    prior(lkj(4), class = cor)),
          data = children, family = bernoulli, sample_prior = TRUE,
          iter = 4000, warmup = 2000, chains = 4, cores = 8,
          control = list(adapt_delta = 0.99))
```

### Check the model fit.
```{r}
plot(m2, ask = FALSE)
```

### Posterior Predictive Check
```{r}
pp_check(m2)
```

### Model Summary
```{r}
summary(m2, prob = .95)
```

### Hypothesis Testing
```{r}
hypotheses2 <- c("zA = 0", # There is no effect of age.
                 "zI = 0", # There is no effect of inhibition.
                 "zSTM = 0", # There is no effect of short term memory.
                 "zWM = 0") # There is no effect of working memory.
hyp2 <- hypothesis(m2, hypothesis = hypotheses2, class = "b")
plot(hyp2)
```

```{r}
hyp2$hypothesis$Evid.Ratio %>% round(digits = 2)

# H0: There is no effect of age, inhibition, short-term memory, working memory, respectively.
# H1: There is an effect of age, inhibition, short-term memory, working memory, respectively.

# Age: Anecdotal evidence for H0
# Inhibition: Anecdotal evidence for H1
# Short-Term: Anecdotal evidence for H0
# Working Memory: Anecdotal evidence for H0
```

### Conditional Effects
```{r}
plot(conditional_effects(m2), ask = FALSE)
```

### Bayesian R^2
```{r}
bayes_R2(m2)
```

### Extract the posterior samples of the parameters
```{r}
post2 <- as_draws_df(m2)
```

### If an effect is negative, check the probability that it is positive; if positive, or vice versa.
```{r}
mean(post2$b_zA < 0)
mean(post2$b_zI > 0)
mean(post2$b_zSTM < 0)
mean(post2$b_zWM < 0)
```

### Session Info
```{r}
sessionInfo()
```
