---
title: "Eye-tracking-viz"
author: "Semih Can Aktepe"
date: "2023-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Set the working directory
```{r}
setwd("/Users/semih/Desktop")
```

### Load the necessary packages
```{r}
library(brms)
library(ggplot2)
library(stringr)
library(posterior)
library(bayestestR)
library(BayesFactor)
library(marginaleffects)
```

### Load the data set
```{r}
et <- read.csv("~/Desktop/eyetracking.csv")
```

### Convert the data types into appropriate formats
```{r}
et$id <- as.factor(et$id)
et$subject <- as.factor(et$subject)
et$group <- as.factor(et$group)
et$item <- as.factor(et$item)
et$condition <- as.factor(et$condition)
et$outcome <- as.factor(et$outcome)
et$itemtype <- as.factor(et$itemtype)
```

## Visualization

### Define the colors
```{r}
# Define custom colors
col1 <- rgb(red = 1, green = 0.647, blue = 0, alpha = 1)
col1p <- rgb(red = 1, green = 0.647, blue = 0, alpha = 0.25)
col1l <- rgb(red = 1, green = 0.647, blue = 0, alpha = 0.025)
col2 <- rgb(red = 0.12, green = 0.56, blue = 1, alpha = 1)
col2p <- rgb(red = 0.12, green = 0.56, blue = 1, alpha = 0.25)
col2l <- rgb(red = 0.12, green = 0.56, blue = 1, alpha = 0.025)
col3 <- rgb(red = 0.196, green = 0.804, blue = 0.196, alpha = 1)
col3p <- rgb(red = 0.196, green = 0.804, blue = 0.196, alpha = 0.25)
col3l <- rgb(red = 0.196, green = 0.804, blue = 0.196, alpha = 0.025)
col4 <- rgb(red = 0.969, green = 0.325, blue = 0.58, alpha = 1)
col4p <- rgb(red = 0.969, green = 0.325, blue = 0.58, alpha = 0.25)
col4l <- rgb(red = 0.969, green = 0.325, blue = 0.58, alpha = 0.025)
```

### Prepare the data for visualization
```{r}
# Binarize the binned data.
et$t <- ifelse(et$target>5, 1, 0)
et$c <- ifelse(et$competitor>5, 1, 0)
et$cluster <- paste(str_sub(et$group, 1, 2), str_sub(et$itemtype), sep = "")
# This is going to be useful for setting up Gaussian process terms.
et$cluster <- as.factor(et$cluster)
# Remove the points that are outside either of the AoIs.
et <- et[et$t==1 | et$c==1,]
```

```{r}
et$ss <- 1 # Dummy variable for standard error and confidence interval calculations.
d1 <- aggregate(t ~ Time + cluster, data = et, FUN = mean)
d2 <- aggregate(t ~ Time + cluster, data = et, FUN = sd)
d3 <- aggregate(ss ~ Time + cluster, data = et, FUN = sum)

d <- cbind(d1, sd = d2$t, ss = d3$ss) #%>% arrange(condition, outcome)
d$se <- d$sd/sqrt(d$ss)
alpha = 0.05
d$ci <- qt(p=alpha/2, df=d$ss,lower.tail=F)*(d$sd/sqrt(d$ss))
d$Low <- d$t - d$ci
d$High <- d$t + d$ci
rm(d1,d2,d3)
```

```{r}
len_d <- nrow(d)
l <- c(1)
for (i in 1:8){
  l[i+1] <- l[i] + len_d/8
}
l
```

#### Print the plot
```{r}
ts <- 0:48*100 #Define a time series
#par(mfrow=c(2,2))
# Counterfactual Negative
plot(NULL, xlim = c(0,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Counterfactual Negative", bty = "l")
polygon(c(ts, rev(ts)), c(d$Low[1:49], rev(d$High[1:49])), col = col1p, border = NA)
lines(ts, d$t[1:49], type = "l", lwd = 2, col = col1)
polygon(c(ts, rev(ts)), c(d$Low[197:245], rev(d$High[197:245])), col = col2p, border = NA)
lines(ts, d$t[197:245], type = "l", lwd = 2, col = col2)
abline(v=c(675, 1241, 2204, 2453, 3806), lty = 3)
abline(h=0.5, lty = 1)
text(c(0, 675, 2204, 2453), 0,
     c("Can", "arabayı", "sürseydi", "işe zamanında yetişecekti."), 
     pos = c(4,4,2,4), font = 3, cex = 0.75)
text(c(0, 675, 2204, 2453, 3806), 1,
     c("John", "car", "had driven", "job on time would've been.", "Target: Bicycle"), 
     pos = c(4,4,2,4,4), font = c(3,3,3,3,2), cex = 0.75)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), 
       col = c(col1, col2), lty = 1, lwd = 2, cex = 0.7)

# Counterfactual Positive
plot(NULL, xlim = c(0,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Counterfactual Positive", bty = "l")
polygon(c(ts, rev(ts)), c(d$Low[50:98], rev(d$High[50:98])), col = col1p, border = NA)
lines(ts, d$t[50:98], type = "l", lwd = 2, col = col1)
polygon(c(ts, rev(ts)), c(d$Low[246:294], rev(d$High[246:294])), col = col2p, border = NA)
lines(ts, d$t[246:294], type = "l", lwd = 2, col = col2)
abline(v=c(675, 1241, 2204, 2453, 3806), lty = 3)
abline(h=0.5, lty = 1)
text(c(0, 675, 2204, 2453, 3806), 0,
     c("Can", "bisikleti", "sürseydi", "işe geç", "kalacaktı."), 
     pos = c(4,4,2,4,2), font = 3, cex = 0.75)
text(c(0, 675, 2204, 2453, 3806, 3806), 1,
     c("John", "bicycle", "had ridden", "job late", "would've been.", "Target: Car"), 
     pos = c(4,4,2,4,2,4), font = c(3,3,3,3,3,2), cex = 0.75)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), 
       col = c(col1, col2), lty = 1, lwd = 2, cex = 0.7)

# Indicative Negative
plot(NULL, xlim = c(0,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Indicative Negative", bty = "l")
polygon(c(ts, rev(ts)), c(d$Low[99:147], rev(d$High[99:147])), col = col1p, border = NA)
lines(ts, d$t[99:147], type = "l", lwd = 2, col = col1)
polygon(c(ts, rev(ts)), c(d$Low[295:343], rev(d$High[295:343])), col = col2p, border = NA)
lines(ts, d$t[295:343], type = "l", lwd = 2, col = col2)
abline(v=c(675, 1241, 2204, 2453, 3806), lty = 3)
abline(h=0.5, lty = 1)
text(c(0, 675, 2204, 2453, 3806), 0,
     c("Can", "bisikleti", "sürerse", "işe geç", "kalacak."), 
     pos = c(4,4,2,4,2), font = 3, cex = 0.75)
text(c(0, 675, 2204, 2453, 3806, 3806), 1,
     c("John", "bicycle", "rides", "job late", "will be.", "Target: Car"), 
     pos = c(4,4,2,4,2,4), font = c(3,3,3,3,3,2), cex = 0.75)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), 
       col = c(col1, col2), lty = 1, lwd = 2, cex = 0.7)

# Indicative Positive
plot(NULL, xlim = c(0,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Indicative Positive", bty = "l")
polygon(c(ts, rev(ts)), c(d$Low[148:196], rev(d$High[148:196])), col = col1p, border = NA)
lines(ts, d$t[148:196], type = "l", lwd = 2, col = col1)
polygon(c(ts, rev(ts)), c(d$Low[344:392], rev(d$High[344:392])), col = col2p, border = NA)
lines(ts, d$t[344:392], type = "l", lwd = 2, col = col2)
abline(v=c(675, 1241, 2204, 2453, 3806), lty = 3)
abline(h=0.5, lty = 1)
text(c(0, 675, 2204, 2453), 0,
     c("Can", "arabayı", "sürerse", "işe zamanında yetişecek."), 
     pos = c(4,4,2,4), font = 3, cex = 0.75)
text(c(0, 675, 2204, 2453, 3806, 3806), 1,
     c("John", "car", "drives", "job on time", "will be.", "Target: Car"), 
     pos = c(4,4,2,4,2,4), font = c(3,3,3,3,3,2), cex = 0.75)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), 
       col = c(col1, col2), lty = 1, lwd = 2, cex = 0.7)

#par(mfrow=c(1,1))
```

## Statistical Analysis

### Selecting the period of interest (PoI).
Our PoI is the time window starting from the offset of the verb in the antecedent to the beginning of the comprehension question.
```{r}
et1 <- et[et$Time>2199,]
```

### Prepare the data for statistical analysis
```{r}
et2 <- aggregate(ss ~ bin + cluster, data = et1, FUN=sum)
et3 <- aggregate(t ~ bin + cluster, data = et1, FUN=sum)
et4 <- aggregate(c ~ bin + cluster, data = et1, FUN=sum)
et2 <- cbind(et2, target = et3$t, competitor = et4$c)
et2$obs <- et2$ss
et2$ss <- NULL
rm(et3,et4)
```

### Fit the model
```{r}
m3 <- brm(target | trials(obs) ~ 1 + cluster + # Intercept and fixed effect of cluster of group, condition, and outcome
            gp(bin, by = cluster), # Gaussian process term varying for each cluster
          prior = c(prior(normal(0, 1), class = Intercept),
                    prior(normal(0, 1), class = b),
                    prior(inv_gamma(3, 1), class = lscale),
                    prior(exponential(1), class = sdgp)),
          data = et2, family = binomial, sample_prior = TRUE,
          iter = 4000, warmup = 2000, chains = 4, cores = 8,
          control = list(adapt_delta = 0.99, max_treedepth = 12))
```

### Check the model fit.
```{r}
plot(m3, ask = FALSE)
```

### Posterior predictive check.
```{r}
pp_check(m3, ask = FALSE)
```

### Model Summary
```{r}
summary(m3, prob = .95)
```

### Leave-One-Out Cross Validation
```{r}
loo(m3, cores = 8)
```

### Bayesian R^2
```{r}
bayes_R2(m3)
```

## Posterior predictives

### Extract posterior samples
```{r}
post_m3 <- as.data.frame(posterior_epred(m3, newdata = et2))
```

#### Get the proportions
```{r}
for (i in 1:216){
  post_m3[,i] <- post_m3[,i]/et2$obs[i]
}
```

### Get the means
```{r}
p_mean <- c()
for (i in 1:216){
  p_mean[i] <- mean(post_m3[,i])
}
```

### Get the 95% CrIs
```{r}
pred <- ci(post_m3, method = "HDI")
pred$est <- p_mean
```

### Plot the posterior predictives
```{r}
# Define bin series
bs <- 22:48*100
# Counterfactual Negative
plot(NULL, xlim = c(2200,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Counterfactual Negative", bty = "l")
polygon(c(bs, rev(bs)), c(pred$CI_low[1:27], rev(pred$CI_high[1:27])), col = col1p, border = NA)
lines(bs, pred$est[1:27], col = col1, lwd = 2)
polygon(c(bs, rev(bs)), c(pred$CI_low[109:135], rev(pred$CI_high[109:135])), col = col2p, border = NA)
lines(bs, pred$est[109:135], col = col2, lwd = 2)
abline(h = 0.5, lty = 2)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), col = c(col1, col2), 
       lty = 1, lwd=2, cex = 0.75)

# Counterfactual Positive
plot(NULL, xlim = c(2200,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Counterfactual Positive", bty = "l")
polygon(c(bs, rev(bs)), c(pred$CI_low[28:54], rev(pred$CI_high[28:54])), col = col1p, border = NA)
lines(bs, pred$est[28:54], col = col1, lwd = 2)
polygon(c(bs, rev(bs)), c(pred$CI_low[136:162], rev(pred$CI_high[136:162])), col = col2p, border = NA)
lines(bs, pred$est[136:162], col = col2, lwd = 2)
abline(h = 0.5, lty = 2)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), col = c(col1, col2), 
       lty = 1, lwd=2, cex = 0.75)

# Indicative Negative
plot(NULL, xlim = c(2200,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Indicative Negative", bty = "l")
polygon(c(bs, rev(bs)), c(pred$CI_low[55:81], rev(pred$CI_high[55:81])), col = col1p, border = NA)
lines(bs, pred$est[55:81], col = col1, lwd = 2)
polygon(c(bs, rev(bs)), c(pred$CI_low[163:189], rev(pred$CI_high[163:189])), col = col2p, border = NA)
lines(bs, pred$est[163:189], col = col2, lwd = 2)
abline(h = 0.5, lty = 2)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), col = c(col1, col2), 
       lty = 1, lwd=2, cex = 0.75)

# Indicative Positive
plot(NULL, xlim = c(2200,4800), ylim = c(0,1), xlab = "Time", ylab = "Target Looks",
     main = "Indicative Positive", bty = "l")
polygon(c(bs, rev(bs)), c(pred$CI_low[82:108], rev(pred$CI_high[82:108])), col = col1p, border = NA)
lines(bs, pred$est[82:108], col = col1, lwd = 2)
polygon(c(bs, rev(bs)), c(pred$CI_low[190:216], rev(pred$CI_high[190:216])), col = col2p, border = NA)
lines(bs, pred$est[190:216], col = col2, lwd = 2)
abline(h = 0.5, lty = 2)
legend("bottomright", inset = 0.03, title = "Age Group", c("Adults", "Children"), col = c(col1, col2), 
       lty = 1, lwd=2, cex = 0.75)
```

### Extract the posterior of the parameters
```{r}
ps <- as.data.frame(as_draws_df(m3))
```

### Covariance Kernel

### Extract the parameters to visualize the covariance kernel
```{r}
sdgp_mean <- c()
lscale_mean <- c()
for (i in 9:16){
  sdgp_mean[i] <- mean(ps[,i])
}
for (i in 17:24){
  lscale_mean[i] <- mean(ps[,i])
}
sdgp_mean <- na.omit(sdgp_mean)
lscale_mean <- na.omit(lscale_mean)
```

### Plot the covariance Kernel for each cluster
```{r}
plot(NULL, xlim = c(1,26), ylim = c(0,1), xlab = "Distance (ms.)", ylab = "Covariance",
     main = "Covariance Kernel", bty = "l", xaxt = "n")
for (i in 1:8){
  curve(sdgp_mean[i]*exp(-(x^2)/2*lscale_mean[i]^2), add = TRUE, lwd = 4, col = i)
}
axis(1, at=c(1,5,10,15,20,25), c(100, 500, 1000, 1500, 2000, 2500))
legend("topright", inset = 0.03, title = "Cluster", levels(et$cluster), col = 1:8, lty = 1, lwd = 4)
```

### Session Info
```{r}
sessionInfo()
```
